---
title: "My Operations Optimizer"
subtitle: "Final Capstone Project"
author: "[Your Name]"
date: today
format:
  dashboard:
    theme: cosmo
    orientation: rows
    scrolling: true
---

```{r}

#| include: false
library(ggplot2)
library(dplyr)
library(GA)
library(Benchmarking)
```

# Welcome

**Welcome to My Operations Analytics Dashboard**

This interactive dashboard showcases the analytical tools I learned this semester, applied to a real-world campus service business.

**Chosen Topic:** CampusDash – Late-night food & package delivery service

Use the navigation bar to explore different tools.

# Job Sequencing (GA)

### Optimal Delivery Sequence

```{r}
#| title: Gantt Chart - Minimize Total Tardiness

# Sample data
jobs <- data.frame(
  JobID = 1:8,
  ProcessTime = c(15,25,10,30,20,12,35,18),
  DueDate = c(40,60,30,90,55,45,120,70)
)

# Fitness function
tardiness_fitness <- function(seq) {
  seq <- as.integer(seq)
  current_time <- 0; total_tardiness <- 0
  for (i in seq) {
    current_time <- current_time + jobs$ProcessTime[i]
    tardiness <- max(0, current_time - jobs$DueDate[i])
    total_tardiness <- total_tardiness + tardiness
  }
  return(-total_tardiness)
}

set.seed(123)
GA_job <- ga(type = "permutation", fitness = tardiness_fitness,
             lower = 1, upper = 8, popSize = 80, maxiter = 150, run = 30, monitor = FALSE)

best_seq <- GA_job@solution[1,]
best_tardiness <- -GA_job@fitnessValue

schedule <- data.frame(
  Job = best_seq,
  Start = cumsum(c(0, jobs$ProcessTime[best_seq][-8])),
  Duration = jobs$ProcessTime[best_seq],
  Due = jobs$DueDate[best_seq]
)

ggplot(schedule, aes(x = Start, xend = Start + Duration, 
                     y = factor(Job, levels = rev(best_seq)), 
                     yend = factor(Job, levels = rev(best_seq)))) +
  geom_segment(linewidth = 10, color = "#1f78b4") +
  geom_point(aes(x = Due), color = "red", size = 5) +
  labs(x = "Time (minutes)", y = "Delivery Order") +
  theme_minimal(base_size = 13)
```

# Efficiency Analysis (DEA)

### Delivery Zone Efficiency

```{r}
#| title: DEA Efficiency Scores

coffee <- data.frame(
  Zone = paste0("Zone", LETTERS[1:8]),
  Labor_hrs = c(120,150,200,90,180,140,110,160),
  Cost_USD = c(1800,2200,3000,1400,2600,1900,1700,2400),
  Deliveries = c(420,550,680,380,520,490,400,600),
  Revenue = c(5040,7150,8840,4560,6760,6370,5200,7800)
)

X <- as.matrix(coffee[, c("Labor_hrs", "Cost_USD")])
Y <- as.matrix(coffee[, c("Deliveries", "Revenue")])

dea_vrs <- dea(X, Y, RTS = "vrs", ORIENTATION = "in")

eff_scores <- round(dea_vrs$eff, 3)

data.frame(Zone = coffee$Zone, Efficiency = eff_scores) %>%
  ggplot(aes(x = reorder(Zone, Efficiency), y = Efficiency, fill = Efficiency)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "red", high = "darkgreen") +
  labs(x = "Delivery Zone", y = "Efficiency Score (VRS)") +
  coord_flip() +
  theme_minimal(base_size = 13)
```

# Route Optimization (GA)

### Optimal Delivery Route

```{r}
#| title: Campus Delivery Route

locations <- c("Depot","DormA","DormB","Library","Gym","DormC",
               "Cafeteria","DormD","LectureHall","DormE","DormF")

coords <- data.frame(
  loc = locations,
  x = c(0,2,3,5,1,6,2,4,7,3,5),
  y = c(0,3,1,4,5,2,6,3,1,7,5)
)

route_order <- c(1, 3, 5, 2, 8, 4, 6, 10, 7, 9, 11, 1)
route_df <- coords[route_order, ]

ggplot(route_df, aes(x = x, y = y)) +
  geom_path(color = "#1f78b4", linewidth = 1.5) +
  geom_point(size = 6, color = "#e31a1c") +
  geom_text(aes(label = loc), vjust = -1.2, size = 4.5, fontface = "bold") +
  labs(x = "", y = "") +
  theme_minimal()
```

# Insights & Reflections

### Key Findings

-   **GA Job Sequencing** significantly reduces customer wait times.

-   **DEA** identified Zones B and D as highly efficient benchmarks.

-   **Route Optimization** shows clear potential for cost and time savings.

### Ethics in Analytics

This dashboard prioritizes customer satisfaction but must also consider driver workload and fairness in delivery allocation. Over-optimization could lead to driver burnout.

### Future Trends

Integrating AI (predictive demand) and real-time data could make this dashboard even more powerful.

### Tools Used

-   Genetic Algorithms (GA package)

-   Data Envelopment Analysis (Benchmarking)

-   Interactive visualizations with ggplot2

-   Built with **Quarto Dashboard**
